Selecting and ranking interesting pairs of cyclic temporal granularities

### 2021-01-13 paper/main.Rmd
update all files to run raw N(0,1) for different designs

### Notations Definitions

@@

# Behavior of raw wpd (weighted pairwise distances)

@@

We now discuss the effect of different tuning parameters $\lambda$ and $\omega$ with the help of two alternate designs ($D_{var_x}$ and $D_{var_f}$). Figure \ref{fig:tuning-one-omega} displays different values of $\lambda$ for a relatively small and higher $\omega$ for $nx = nfacet =  \{2, 3, 5, 7, 9\}$. The $\lambda$ for which the two designs intersect will then be chosen as the optimal $\lambda$ that could then be weighed appropriately to up-weigh the within-facet distances and down-weigh the between-facet distances. The design $D_{var_x}$ increases linearly with increasing $\lambda$, whereas $D_{var_f}$ decreases linearly with increasing $\lambda$. This is expected as wpd has a linear relationship with $\lambda$ by construction. The two designs mostly intersect at $\lambda = 0.5$ for a higher value of $\omega$.  \autoref{fig:tuning-parameter-intersection} shows how the value of $\lambda$ changes with increasing $\omega$ for $\omega =  \{1, 2, \dots, 10\}$ and $\lambda = seq(from  = 0.1, to = 0.9, by = 0.05)$.

@@

```{r tuning-one-omega, fig.caption = "WPD from two designs are plotted for different values of lambda for different x and facet categories. For omega = 1, the designs intersect for 0.6<lamba<=0.75, whereas for higher omega, design intersects at lambda = 0.5"}
knitr::include_graphics(here::here("simulations/tuning_param/figs/fixed_omega.png"))
```

```{r tuning-parameter-intersection, fig.cap="For most panels it is observed that the most common value of the tuning paramter for which the designs interact is 0.5, which implies any value greater than 0.5 could be chosen to up-weigh the within-facet distances and down-weigh the between-facet distances for most sitatuions."}
knitr::include_graphics(here::here("simulations/tuning_param/figs/intersection_plot.png"))
```

How value of lambda changes with increasing omega and mean for fixed nx, nfacet and standard deviation?

```{r lambda-omega-mean, eval = FALSE}
knitr::include_graphics(here::here("simulations/result_report/lambda_omega_mean_3_2.png"))
```

How value of lambda changes with increasing omega and sd for fixed nx, nfacet and mean?

```{r lambda-omega-sd, eval = FALSE}
knitr::include_graphics(here::here("simulations/result_report/lambda_omega_sd_3_2.png"))
```

```{r lamba-omega-alternate-design}
knitr::include_graphics((here::here("simulations/result_report/wpd_alterate_designs.png")))
```

 Distribution across facet and x categories
<!-- ### Need for normalisation -->

From Figure \ref{fig:dist-across-facets} and \ref{fig:dist-across-x} shows the MMPD distribution is different for different levels of facets and x-axis levels. With increasing number of facets, the location of the distribution shift rightwards and with increasing x-axis levels, the scale of the distribution reduces.

Median and maximum distances are affected by the number of categories considered and hence the distance measure $MMPD_{raw}$, which is a combination of median and maximum would also be influenced by the number of levels. It would have higher values if $C_i$ or $C_j$ has higher levels. We would ideally want a higher value of the measure if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories are higher. In Figures \ref{fig:dist-across-facets} and \ref{fig:dist-across-x}, both the cyclic granularities $C_i$ and $C_j$ are considered such that their levels vary in the range $(2, 5,7 ,9, 14, 21, 30, 45)$. Each of these combinations is considered a panel for each of which $MMPD_{raw}$ has been constructed $100$ times to compute the distribution. <!--More specifically, Figure \ref{fig:mean-sd-max} shows the marginal contribution of increasing levels of $C_i$ and $C_j$ on distribution of $MMPD_raw$. This shows that with the increase in the number of levels of $C_i$, the location of the distribution shift rightward, whereas for increasing x-axis levels, the scale of the distribution decreases.-->

Therefore, in order to compare $MMPD_{raw}$ across different combinations of facet and x-axis levels, we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure.

@@

## Sample size
## Number of permutations

@@

In an attempt to make the densities \ref{fig:hist_plot} same across different levels, we want to start by making their scale and location same. In that regard, we shuffle the data multiple times and compute the sampling distribution of $mmpd_{raw}$ for each combination of facet and x-axis levels.

@@

## Simulation study

@@

```{r}
knitr::include_graphics(here::here("paper/Figs/simulation.png"))
```

@@

## Results

@@

```{r norm-data, fig.cap = "Distribution of $MMPD_{norm}$ across different levels of facet and x-axis. These distributions should look the same if normalisation worked." }
mmpd_norm = read_rds("../hakear-drake/data/data_norm_low_new.rds") 
hist_plot = mmpd_norm %>% 
      ggplot() +
      geom_histogram(aes(x = raw_mmpd)) +
      facet_grid(nx~nfacet)
hist_plot
```

```{r null-dist, eval =  FALSE, echo = FALSE, fig.cap="Distribution of mmpd across different facet levels (columns) and x-axis levels (rows) for null ditribution normal(5,10) and gamma(0.5, 0.2). The distribution of MMPD is unaffected by the null distribution considered. It is also the same across all facet levels but not across the x-axis, which means normalisation is failing across x-axis."}
# all_files = list.files(path = "../hakear-drake/data", pattern = ".rds")
# all_files
# 
# read_files = lapply(all_files, function(x){
#   read_rds(paste0("../hakear-drake/data/", x))
# })
# 
# names(read_files) = str_remove(all_files, pattern = ".rds")
# 
# null_all = bind_rows(normal = read_files$null_normal,
#                      exponential = read_files$null_exp,
#                      gamma = read_files$null_gamma,
#                      weibull = read_files$null_weibull,
#                      .id = "distribution")
# design_all = readr::read_rds("../hakear-drake/data/design_all.rds")
design_all = readr::read_rds("data/design_all.rds")

@@

### Performance of normalised MMPD

```{r}
data1  = sim_panel(nx = 2, nfacet = 3, ntimes = 500)
simulated_data <- sim_panel(nx = 2,
nfacet = 3,
ntimes = 500,
sim_dist = rep(dist_normal(seq(5, 15, 5), 5), each = 2))
data2 = simulated_data
data4 = sim_panel(nx = 2,
nfacet = 3,
ntimes = 500,
sim_dist = dist_normal(seq(5,30, 5), 5))
data3 = data2 %>% rename("temp" = "id_facet") %>% rename("id_facet" = "id_x") %>% rename("id_x" = "temp")

>>>>

### Notations and Definitions

@@

# Behavior of raw wpd

@@

We now discuss the effect of different tuning parameters $\lambda$  with the help of two alternate designs ($D_{var_x}$ and $D_{var_f}$) and $\omega$. \autoref{fig:tuning-oneomega} displays different values of $\lambda$ for a relatively small and higher $\omega$ for $nx = nfacet =  \{2, 3, 5, 7, 9\}$. The $\lambda$ for which the two designs intersect will then be chosen as the optimal $\lambda$ that could then be weighed appropriately to up-weigh the within-facet distances and down-weigh the between-facet distances. The design $D_{var_x}$ increases linearly with increasing $\lambda$, whereas $D_{var_f}$ decreases linearly with increasing $\lambda$. This is expected as wpd has a linear relationship with $\lambda$ by construction. The two designs mostly intersect at $\lambda = 0.5$ for a higher value of $\omega$.  \autoref{fig:tuning-parameter-inter} shows how the value of $\lambda$ changes with increasing $\omega$ for $\omega =  \{1, 2, \dots, 10\}$ and $\lambda = seq(from  = 0.1, to = 0.9, by = 0.05)$.

@@

```{r tuning-oneomega, fig.caption = "WPD from two designs are plotted for different values of lambda for different x and facet categories. For omega = 1, the designs intersect for 0.6<lamba<=0.75, whereas for higher omega, design intersects at lambda = 0.5"}
knitr::include_graphics(here::here("simulations/tuning_param/figs/fixed_omega.png"))
```

```{r tuning-parameter-inter, fig.cap ="For most panels it is observed that the most common value of the tuning paramter for which the designs interact is 0.5, which implies any value greater than 0.5 could be chosen to up-weigh the within-facet distances and down-weigh the between-facet distances for most sitatuions."}
knitr::include_graphics(here::here("simulations/tuning_param/figs/intersection_plot.png"))
```

We now discuss the values wpd takes for different designs. Ideally the wpd should have the higherst values for $D_{var_{all}}$, followed by $D_{var_x}$ and $D_{var_y}$ for $\lambda \geq 0.5$. We test this for lower and higher values of $\omega = c(1, 8, 15, 50)$

<!-- How value of lambda changes with increasing omega and mean for fixed nx, nfacet and standard deviation? -->

<!-- # ```{r lambda-omega-mean, eval = FALSE} -->
<!-- # knitr::include_graphics(here::here("simulations/result_report/lambda_omega_mean_3_2.png")) -->
<!-- # ``` -->
<!-- #  -->
<!-- # How value of lambda changes with increasing omega and sd for fixed nx, nfacet and mean? -->
<!-- #  -->
<!-- # ```{r lambda-omega-sd, eval = FALSE} -->
<!-- # knitr::include_graphics(here::here("simulations/result_report/lambda_omega_sd_3_2.png")) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r lamba-omega-alternate-design} -->
<!-- # knitr::include_graphics((here::here("simulations/result_report/wpd_alterate_designs.png"))) -->
<!-- # ``` -->

 <!-- Distribution across facet and x categories -->
<!-- ### Need for normalisation -->

### Discussion

The distribution of wpd is different for different levels of facets and x-axis levels. This is because the statistic maximum which is used to define wpd is affected by the number of categories. The measure would have higher values if $C_i$ or $C_j$ has higher levels. However, we would ideally want a higher value of the measure only if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories are higher. Therefore, in order to compare wpd across different combinations of facet and x-axis levels, we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure. Henceforth we call the measure discussed as $wpd_{raw}$ and the normalised measure as $wpd_{norm}$. The measure $wpd_{norm}$ could potentially lead to comparison of the measure across different panels and also identifying only the interesting panels from a dataset.

<!-- In Figures \ref{fig:dist-across-facets} and \ref{fig:dist-across-x}, both the cyclic granularities $C_i$ and $C_j$ are considered such that their levels vary in the range $(2, 5,7 ,9, 14, 21, 30, 45)$. Each of these combinations is considered a panel for each of which $MMPD_{raw}$ has been constructed $100$ times to compute the distribution. <!--More specifically, Figure \ref{fig:mean-sd-max} shows the marginal contribution of increasing levels of $C_i$ and $C_j$ on distribution of $MMPD_raw$. This shows that with the increase in the number of levels of $C_i$, the location of the distribution shift rightward, whereas for increasing x-axis levels, the scale of the distribution decreases.-->

@@

@@

In an attempt to make the densities \ref{fig:hist_plot} same across different levels, we want to start by making their scale and location same. In that regard, we shuffle the data multiple times and compute the sampling distribution of  for each combination of facet and x-axis levels.

@@

@@

```{r, eval = FALSE}
knitr::include_graphics(here::here("paper/Figs/simulation.png"))
```

@@

## Sample size

How smaller or different sample size play a role in the distribution of the measure.

## Number of permutations

Find the number of permutations which are sufficient for consistent estimation of mean and standard deviation

## Discussion

Did normalisation work?

@@

```{r norm-data, fig.cap = "Distribution of $MMPD_{norm}$ across different levels of facet and x-axis. These distributions should look the same if normalisation worked." , eval = F}
mmpd_norm = read_rds("../hakear-drake/data/data_norm_low_new.rds") 
hist_plot = mmpd_norm %>% 
      ggplot() +
      geom_histogram(aes(x = raw_mmpd)) +
      facet_grid(nx~nfacet)
hist_plot
```

```{r null-dist, eval =  FALSE, echo = FALSE, fig.cap="Distribution of mmpd across different facet levels (columns) and x-axis levels (rows) for null ditribution normal(5,10) and gamma(0.5, 0.2). The distribution of MMPD is unaffected by the null distribution considered. It is also the same across all facet levels but not across the x-axis, which means normalisation is failing across x-axis.", eval = F}
# all_files = list.files(path = "../hakear-drake/data", pattern = ".rds")
# all_files
# 
# read_files = lapply(all_files, function(x){
#   read_rds(paste0("../hakear-drake/data/", x))
# })
# 
# names(read_files) = str_remove(all_files, pattern = ".rds")
# 
# null_all = bind_rows(normal = read_files$null_normal,
#                      exponential = read_files$null_exp,
#                      gamma = read_files$null_gamma,
#                      weibull = read_files$null_weibull,
#                      .id = "distribution")
# design_all = readr::read_rds("../hakear-drake/data/design_all.rds")
design_all = readr::read_rds("data/design_all.rds")

@@

<!-- ### Performance of normalised MMPD -->

```{r, eval = FALSE}
data1  = sim_panel(nx = 2, nfacet = 3, ntimes = 500)
simulated_data <- sim_panel(nx = 2,
nfacet = 3,
ntimes = 500,
sim_dist = rep(dist_normal(seq(5, 15, 5), 5), each = 2))
data2 = simulated_data
data4 = sim_panel(nx = 2,
nfacet = 3,
ntimes = 500,
sim_dist = dist_normal(seq(5,30, 5), 5))
data3 = data2 %>% rename("temp" = "id_facet") %>% rename("id_facet" = "id_x") %>% rename("id_x" = "temp")
